---
title: "Project Two"
output: Cierra Disedare [project_two]
  pdf_document: default
  html_document: default
---

Due Oct. 21 at 11:59 PM. 

```
download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2022/main/vignettes/project_two.Rmd", destfile = "/cloud/project/projects/project_2/project_two.Rmd")
```

For this first part of the exam, you can either use `surveys_complete.csv` or your own data. If you are using your own data, you must have data in which you think you have a numerical predictor variable and a numerical response variable. If you are using `surveys_complete`, you can use weight and hindfoot_length for this.

Save this file in your `projects` directory. You can either save it in a project two subdirectory, or just put it in projects. Either way is fine.


1) Load in your data. Which variable will you be using as a predictor, and which as a response? (5 pts)

```{r}
read.csv("/cloud/project/data/surveys.csv")
str(surveys_complete)
na.omit(surveys_complete)
```

```
# The width column is the predictor (x) and the hindfoot_length is the response (y).
```

2) Plot the two against each other with a scatter plot. Do the data appear to be related linearly? (5 pts)


```{r}
scat_pro_2 <- ggplot(data = surveys_complete, mapping = aes (x = weight, y = hindfoot_length)) + geom_point() + theme(text = element_text(size = 16), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) + geom_jitter()
scat_pro_2
```

```
#Yes, the data does appear to be related linearly.
```


3) Fit the linear model. View the summary. (5 pts)


```{r}
model_fit <- lm(hindfoot_length ~weight, data = surveys_complete)
summary(model_fit)
```

4) Does the summary make sense when you compare it to the plot you made? Does our model have good predictive power? Evaluate the residual standard error, intercept, and R-Squared in particular. Would you say your predictor predicts the response?  (10 pts)


```
# Yes, most of the plots points line up with and correlate to the model fit summary. R-squared is less than 0.50 and is non-linear, my intercept is positive and over 20, the opposite of the expected of a linear intercept. Yes, I would say the predictor predicts the response.
```


5) Plot the model on the graph. Increase the size of the text so it is comfortably readable at 5 feet. Make sure axis labels are readable and not overlapping with one another. (5 pts)

```
ggplot(surveys_complete, aes(x = weight, y = hindfoot_length)) + geom_point() + scale_color_manual(values = c("blue", "orange")) + geom_smooth(method = "lm") + annotate("text", x = 40, y = 70, label = "R^2 == 0.988", parse = T, size = 5)
```


6) Check the normality of the residuals. Do they look OK? Are we violating assumptions? (5 pts)

```{r}
qqnorm(augmented_fit$.resid)
qqline(augmented_fit$.resid, col = "deeppink")

# The points look mostly linear with the line.
```

Why is normality of residuals important? 

```{r}

#It is important to ensure the data fits the model used.
```

7) If you are using `surveys_complete`: Is there interspecific variation in the linear model? Hint: look at our prior work with facet plots. (15 pts) 

If you are *not* using  `surveys_complete`: Do you think there are groupings within your data that may have a separate linear model? Try grouping the data by that variable and redoing the lm. If this would not make sense for your data, you may also attempt to do multiple predictor variables. (15 pts)

```{r}
#Yes there is interspecific variation in the linear model because the predictor and response may have small to large differences between species within the model.
```

## Part Two

In this portion, you are free to use your own data if you have a categorical predictor variable and a response variable. Otherwise, you may use the columns sex and weight in `surveys_complete`

1) First, plot the data grouped by sex (5 pts)

```{r}
ggplot(data = surveys_complete, mapping = aes (x = sex, y = weight, color = sex)) + geom_point()
```

2) Try an ANOVA of this data (5 pt)

```{r}
ggplot(surveys_complete, aes(x = sex, y = weight, color = sex)) + geom_jitter() + labs(x = "Sex", y = "Weight") + stat_summary(fun.data = "mean_se")

aov(model_fit) -> anova_model_fit

summary(anova_model_fit)
```
>> summary(anova_model_fit)
               Df  Sum Sq Mean Sq F value Pr(>F)    
weight          1 1308874 1308874   26995 <2e-16 ***
Residuals   30736 1490259      48                   
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

3) How about a linear model? What information does this give you that an ANOVA can't? (5 pts)

```{r}
model_fit <- lm(weight ~sex, data = surveys_complete)
summary(model_fit)
```
>> summary(model_fit)

Call:
lm(formula = weight ~ sex, data = surveys_complete)

Residuals:
   Min     1Q Median     3Q    Max 
-38.08 -22.08  -5.47   5.53 237.92 

Coefficients:
            Estimate Std. Error t value Pr(>|t|)    
(Intercept)  41.4701     0.2954 140.391   <2e-16 ***
sexM          0.6124     0.4078   1.502    0.133    
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 35.67 on 30674 degrees of freedom
  (62 observations deleted due to missingness)
Multiple R-squared:  7.351e-05,	Adjusted R-squared:  4.091e-05 
F-statistic: 2.255 on 1 and 30674 DF,  p-value: 0.1332

```
#ANOVA gives risiduals and significant codes only, while linear moel gives those and R-squared, intercepts, and estimated standard error.
```

3) Plot the lm with the data, with points colored by sex. (10 pts)

```{r}
 ggplot(surveys_complete, aes(x = sex, y = weight, color = sex)) + geom_point() + scale_color_manual(values = c("purple", "red")) + geom_smooth(method = "lm") + annotate("text", x = "NA", y = 250, label = "R^2 == 0.988", parse = T, size = 5)
```

4) Choose any model we've looked at so far, but use two predictor variables. Perform an LM, plot the results, and state if this changes your interpretation of the relationship between variables (10 pts)

```{r}
model_fit <- lm(hindfoot_length ~sex + weight, data = surveys_complete)
summary(model_fit)
```

```{r}
> ggplot(surveys_complete, aes(x = weight, y = hindfoot_length, color = sex)) + geom_point() + geom_smooth(method = "lm") + annotate("text", x = 40, y = 30, label = "R^2 == 0.988", parse = T, size = 5)
```

```
# This shows that hindfoot_length is within a much smaller range than weight for the sexes.
```

## Part Three


1) Add and commit this document (5 pts)

```
library(gitcreds)
gitcreds_set()
```

2) Push your changes to github (10 pts)

```
In the gIT TAB, SELECT
```



# MS students

My expectation is that you'll do this with your own data. If any part of this doesn't make sense with your data, please get in touch ASAP so we can work it out.
